\LARGE \bf Informed sampling by Markov Chain Monte Carlo methods

### 2017-02-06 paper.tex
rewrite mcmc

Markov Chain Monte Carlo (MCMC)~\cite{ADDJ03} is a framework of sampling from a given probability distribution by constructing a Markov chain of random transition toward a desired distribution $ \pi (x) $.
A set of samples are generated to approximate the desired distribution $ \pi (x) $.
A approximated desired distribution $ \hat{\pi} (x) $ can be used in any problem that states are modeled random variable, e.g. motion tracking~\cite{KBD04}, data regression~\cite{TL11} and state estimation~\cite{ASC13}.
Unlike a rejection sampling, MCMC does a random walk in a parametric space and concentrates on an important area, which usually brings better efficiency.
Hamiltonian Monte Carlo(HMC)~\cite{N11} is introduced to provide better efficiency in converging to a desired distribution.
It adopts physical system dynamics to propose future states in the Markov chain.
The concept origins from Hamiltonian dynamics, which models the movement of an object, the position $ x $ and the momentum $ q $ over time $ t $.
Define a system energy $ H $ of an object that consists of kinetic energy $ K $ and potential energy $ U $, which is written as $ H(x,q) = U(x) + K(q) $.
The kinetic energy $ K(q) $ is determined by the momentum $ q $ of the object, while the potential energy $ U(x) $ is determined by the position $ x $ of the object.
By the Hamiltonian equation, the movement of the object can be predicted by
$ \frac{dx}{dt} = \frac{\partial H}{\partial q} $ and $ \frac{dq}{dt} = - \frac{\partial H}{\partial x} $.
This provides an approach of the gradient descent by following the decrease of an associated potential energy.
Thus, HMC simulates the Hamiltonian dynamics that moves samples towards a target distribution while avoiding random walks.

A canonical distribution bridges an energy with a probability distribution.
If two different states have the same energy, the probabilities of them are equivalent.
A set of samples from such a distribution can be equivalent to a set of samples driven by an associated energy.
By using the system energy $ H(x, q) $, we can defines a canonical distribution $ P(x, q) = \frac{1}{Z} \exp ( -H(x,q) / T ) $.
We can sample $ [ x ,q ] $ from $ P(x, q) $ and discard the momentum $ q $ (as an auxiliary variable) in order to sample from a marginal distribution $ P(x) $.
If we have a target distribution $ \pi(x) $ of state $ x $ and define a potential energy in the form $  U(x) = - \log ( \pi(x) T ) $, we can have the canonical distribution equal to $ \pi(x) $.
The collected set of $ x $ is equivalent to sampling from $ \pi(x) $.
Thus, we can use simulating the Hamiltonian dynamics to generate a set of samples from a target distribution $ \pi (x) $.
The properties of Hamiltonian dynamics guarantees so that the proposals will always be accepted, which accelerates the convergence rate.

>>>>

An alternative sampling framework, which is widely used in statistical machine learning, is Monte Carlo Simulation~\cite{M97}.
It uses the numerical experimental power of computers to generate samples toward a desired distribution $ \pi (x) $.
The generated set of samples are a discrete approximation $ \hat{\pi} (x) $ of the desired distribution $ \pi (x) $.
A approximated desired distribution $ \hat{\pi} (x) $ can be used in any problem that states are modeled random variable, e.g. motion tracking~\cite{KBD04}, data regression~\cite{TL11} and state estimation~\cite{ASC13}.

Markov Chain Monte Carlo (MCMC)~\cite{ADDJ03} is one most popular class that constructs a Markov chain of random transition that moves samples from an original proposed distribution to a desired distribution $ \pi (x) $.
Unlike a rejection sampling, MCMC does a random walk in a parametric space and concentrates on an important area.
Though the convergence to the desired distribution is proved to be guaranteed, the nature of random walk determines that the convergence of MCMC can be slow in a high-dimensional space.
Hamiltonian Monte Carlo(HMC)~\cite{N11} is thus introduced to reduce the random walk but directly moving to the target by heuristic.

HMC adopts physical system dynamics to propose future states in the Markov chain~\cite{N11}.
It is named ``Hamiltonian, because the concept origins from Hamiltonian dynamics, which associates the system energy with the resulting dynamics.
Define the system dynamics of an object as the position $ x $ and the momentum $ q $ over time $ t $.
We can have a system energy $ H $ of the object that consists of kinetic energy $ K $ and potential energy $ U $, which is written as $ H(x,q) = U(x) + K(q) $.
The kinetic energy $ K(q) $ is determined by the momentum $ q $ of the object, while the potential energy $ U(x) $ is determined by the position $ x $ of the object.
The Hamiltonian dynamics is determined by the gradients of the system energy, which is
$ \frac{dx}{dt} = \frac{\partial H}{\partial q} $ and $ \frac{dq}{dt} = - \frac{\partial H}{\partial x} $.
In HMC, the state to be sampled is used as the position $ x $, and the associated probability $ P(x) $ is defined as the kinematic energy $ K(x) $.
An auxiliary variable is used as the momentum $ q $, which is associated with another probability that is defined as the potential energy $ U(q) $.
Thus, the system energy $ H(x, q) $ is associated with a joint probability $ P(x, q) $. 
As we are interested with sampling from marginal probability $ P(x) $ is interested, we will sample $ [ x ,q ] $ from $ P(x, q) $ and discard the momentum $ q $ as an indirect way of sampling.

The canonical distribution is often choose to bridge a system energy with a probability distribution, which is written as $ P(x, q) = \frac{1}{Z} \exp ( -H(x,q) / T ) $.
If two different states have the same energy, the probabilities of them are equivalent.
Given a target distribution $ \pi(x) $, the potential energy is defined in the form $  U(x) = - \log ( \pi(x) T ) $.

If we have of state $ x $ and define a potential energy , we can have the canonical distribution equal to $ \pi(x) $.
The collected set of $ x $ is equivalent to sampling from $ \pi(x) $.
Thus, we can use simulating the Hamiltonian dynamics to generate a set of samples from a target distribution $ \pi (x) $.
The properties of Hamiltonian dynamics guarantees so that the proposals will always be accepted, which accelerates the convergence rate.
